---
title: "Midterm"
author: "Wenyan Ma (wm2460)"
output: pdf_document
---

# Set up

## Load libraries and data
```{r warning=FALSE, message=FALSE}
library(caret)
library(mgcv)
library(earth)
library(tidyverse)
library(summarytools)
library(corrplot)
library(ggpubr)

setwd("D:/CUMC/Y2S2/DS2/Midterm")

load("./recovery.RData") 

set.seed(2460) 

dat <- dat[sample(1:10000, 2000),]
```

# Exploratory analysis and data visualization

## Data Partition

Here, we mainly want to investigate the EDA of the training dataset. Therefore, we will start with the data partition. 

```{r warning=FALSE, message=FALSE}
set.seed(2460)

trainRows <- createDataPartition(y = dat$recovery_time, p = 0.8, list = FALSE)
```


## Understanding the outcome variable `recovery_time`
```{r warning=FALSE, message=FALSE}
# check the outcome variable
hist(dat$recovery_time[trainRows], breaks = 50)
```

The distribution of the outcome variable `recovery_time` is heavily right-skewed. To account for this, I will take the log-transformation of the outcome and use that variable for following analyses.


```{r warning=FALSE, message=FALSE}
dat = dat %>% 
  na.omit(dat) %>% 
  mutate(lrecovery_time = log(recovery_time)) %>% 
  select(-recovery_time, -id) 
  
# log-transformation helped with making it more normal
hist(dat$lrecovery_time[trainRows], breaks = 50)
```


## Summary of the dataset
```{r warning=FALSE, message=FALSE}
st_options(plain.ascii = F,
           style = "rmarkdown",
           dfSummary.silent = T,
           footnote = NA,
           subtitle.emphasis = F)

dfSummary(dat[trainRows, -1])
```


## Understand categorical variables
```{r warning=FALSE, message=FALSE}
gender = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = gender)) +  geom_bar() +  labs(x = "Gender", y = "Count")

race = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = race)) + geom_bar() + labs(x = "Race",y = "Count")

smoking = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = smoking)) + geom_bar() + labs(x = "Smoking", y = "Count")

hypertension = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = hypertension)) + geom_bar() + labs(x = "Hypertension", 
                                                    y = "Count")
diabetes = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = diabetes)) + geom_bar() + labs(x = "Diabetes",y = "Count")

vaccine = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = vaccine)) + geom_bar() + labs(x = "Vaccination status",
                                               y = "Count")
severity = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = severity)) + geom_bar() + labs(x = "Severity", y = "Count")

study = (dat[trainRows, -1]) %>% 
  ggplot(aes(x = study)) + geom_bar() + labs(x = "Study Site", y = "Count")

cat_combined_plot = ggarrange(gender, race, smoking, hypertension, 
                               diabetes, vaccine,severity, study,
                          ncol = 2, nrow = 4)

cat_combined_plot
```


## Understand continuous variables
```{r warning=FALSE, message=FALSE}
par(mar = c(3, 3, 2, 2), mfrow = c(2, 3))

age = hist(dat$age[trainRows], breaks = 50)

bmi = hist(dat$bmi[trainRows], breaks = 50)

height = hist(dat$height[trainRows], breaks = 50)

weight = hist(dat$weight[trainRows], breaks = 50)

SBP = hist(dat$SBP[trainRows], breaks = 50)

LDL = hist(dat$LDL[trainRows], breaks = 50)
```


## Understand the correlation between predictors
```{r warning=FALSE, message=FALSE}
correlation <- model.matrix(lrecovery_time ~ ., dat)[trainRows,-1]

corrplot(cor(correlation), method = "circle", type = "full")
```


## Understand the relationship with continuous predictors and the outcome
```{r warning=FALSE, message=FALSE}
theme1 <- trellis.par.get()
theme1$plot.symbol$col <- rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch <- 16
theme1$plot.line$col <- rgb(.8, .1, .1, 1)
theme1$plot.line$lwd <- 2
theme1$strip.background$col <- rgb(.0, .2, .6, .2)
trellis.par.set(theme1)

# plotting continuous predictors 
featurePlot(x = model.matrix(lrecovery_time ~ ., dat)[trainRows,c("age", "SBP", "LDL", "height", "weight", "bmi")],
            y = dat$lrecovery_time[trainRows],
            plot = "scatter",
            span = .5,
            labels = c("Predictors","Y"),
            type = c("p", "smooth"),
            layout = c(3,2))
```


## Considering variables based on the EDA

From the correlation plot, we can observe that `bmi` is highly correlated with `weight` and `height`, which makes sense because BMI is calculated by weight divided by the square of height. This demonstrates collinearity between the variables, and to account for this, I will remove the `bmi` variable for the predictions. 

Also, I believe that the `study` variable is more of a geographical indicator to distinguish different study sites, and it will not be critical in predicting recovery time. Therefore, I will also remove  the `study` variable. 

Lastly, I will remove variables `race` and `smoking` since I have created dummy variables for them and I will use the dummy variables in further analyses.

```{r warning=FALSE, message=FALSE}
midterm = dat %>% 
  mutate(
  # create dummy variables for categorical variables
    # set up 3 dummy variables for `race`, reference = White:
         race_2 = ifelse(race == 2, 1, 0), 
         race_3 = ifelse(race == 3, 1, 0),
         race_4 = ifelse(race == 4, 1, 0),
    # set up 2 dummy variables for `smoking`, reference = Never smoked:
         smoking_1 = ifelse(smoking == 1, 1, 0), 
         smoking_2 = ifelse(smoking == 2, 1, 0)) 

# remove variables that will not be used
midterm = midterm %>% 
  select(-bmi, -study, -race, -smoking) 

# partition again based on the new outcome variable
set.seed(2460)

trainRows_new <- createDataPartition(y = midterm$lrecovery_time, p = 0.8, list = FALSE)

x <- model.matrix(lrecovery_time ~ ., midterm)[trainRows_new,-1]

y <- midterm$lrecovery_time[trainRows_new]

x2 <- model.matrix(lrecovery_time ~ ., midterm)[-trainRows_new,-1]

y2 <- midterm$lrecovery_time[-trainRows_new]

ctrl1 <- trainControl(method = "repeatedcv", number = 10, repeats = 5)
```


# Model training

## Mode 1: linear model

### Train the model
```{r warning=FALSE, message=FALSE}
set.seed(2460)

lm.fit <- train(x, y, 
                method = "glm", 
                preProcess = c("center", "scale"),
                trControl = ctrl1)

summary(lm.fit)
```


## Model 2: Ridge

### Train the model
```{r warning=FALSE, message=FALSE}
set.seed(2460)

ridge.fit <- train(x, y,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 0,
                                          lambda = exp(seq(-1, -7, length=100))),
                   preProc = c("center", "scale"),
                   trControl = ctrl1)

ridge.fit$bestTune

plot(ridge.fit, xTrans = log)

coef(ridge.fit$finalModel, s = ridge.fit$bestTune$lambda)
```


## Model 3: Lasso
```{r warning=FALSE, message=FALSE}
set.seed(2460)

lasso.fit <- train(x, y, 
                   method = "glmnet", 
                   tuneGrid = expand.grid(alpha = 1, 
                                          lambda = exp(seq(-9, -3, length = 100))),
                   preProcess = c("center", "scale"),
                   trControl = ctrl1)

lasso.fit$bestTune

plot(lasso.fit, xTrans = log)

coef(lasso.fit$finalModel, s = lasso.fit$bestTune$lambda)
```


## Model 4: Elastic net
```{r warning=FALSE, message=FALSE}
set.seed(2460)

enet.fit <- train(x, y, 
                  method = "glmnet", 
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 21), 
                                         lambda = exp(seq(-8, 1, length = 100))), 
                  preProcess = c("center", "scale"),
                  trControl = ctrl1)

enet.fit$bestTune

myCol <- rainbow(25)

myPar <- list(superpose.symbol = list(col = myCol),
              superpose.line = list(col = myCol))

plot(enet.fit, par.settings = myPar, xTrans = log)

coef(enet.fit$finalModel, s = enet.fit$bestTune$lambda)
```


## Model 5: Partial least square
```{r warning=FALSE, message=FALSE}
set.seed(2460)

pls.fit <- train(x, y,
                 method = "pls",
                 tuneGrid = data.frame(ncomp = 1:15),
                 trControl = ctrl1,
                 preProcess = c("center", "scale"))

pls.fit$bestTune

ggplot(pls.fit, highlight = TRUE) + theme_bw()

summary(pls.fit)

coef(pls.fit$finalModel)
```


## Model 6: GAM
```{r warning=FALSE, message=FALSE}
set.seed(2460)

gam.fit <- train(x, y,
                 method = "gam",
                 tuneGrid = data.frame(method = "GCV.Cp", select = c(TRUE,FALSE)),
                 preProcess = c("center", "scale"),
                 trControl = ctrl1)

gam.fit$bestTune

gam.fit$finalModel

summary(gam.fit)
```


### Plot continuous predictors in GAM
```{r warning=FALSE, message=FALSE}
var.names <- c("age", "SBP", "LDL", "height", "weight", "-")

# make a matrix for easier comprehension of the plot with 16 predictors
matrix <- matrix(var.names, nrow = 2, ncol = 3, byrow = TRUE)

# use the matrix to correspond each plot with each predictor
print(matrix)

gam.plot <- gam.fit$finalModel

# make 16 plots into one
par(mar = c(3, 3, 2, 2), mfrow = c(2, 3))

plot(gam.plot)

title(main = "Predictors' Plot", cex.main = 1, font.main = 3, outer = TRUE, line = -1)
```


## Model 7: MARS
```{r warning=FALSE, message=FALSE}
mars_grid <- expand.grid(degree = 1:3,
                         nprune = 2:25) 

set.seed(2460)

mars.fit <- train(x, y,
                  method = "earth",
                  tuneGrid = mars_grid,
                  preProcess = c("center", "scale"),
                  trControl = ctrl1)

ggplot(mars.fit, highlight = TRUE)

mars.fit$bestTune

coef(mars.fit$finalModel)
```

# Model comparison
```{r warning=FALSE, message=FALSE}
res <- resamples(list(lm = lm.fit, 
                      ridge = ridge.fit,
                      lasso = lasso.fit,
                      enet = enet.fit,
                      pls = pls.fit,
                      gam = gam.fit,
                      mars = mars.fit))

summary(res)

bwplot(res, metric = "RMSE")
```

**Select MARS model as the final model**

## Test error
```{r warning=FALSE, message=FALSE}
mars.pred <- predict(mars.fit, newdata = x2)

mars.testerror = mean((mars.pred - y2)^2)

mars.testerror
```


### Training error
```{r warning=FALSE, message=FALSE}
mars.pred_train <- predict(mars.fit, newdata = x)

mars.trainerror = mean((mars.pred_train - y)^2)

mars.trainerror
```

